---
title: "Gov 50 Final Project"
author: Darcy Lin
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction
Does increasing campaign ads increase votes? In this study, I plan to examine the extent to which campaign ads can predict the 2016 election outcome. I will do so by comparing the proportion of campaign ads that belong to a politician and the proportion of the vote that they received in the 2016 presidential election for that district. 

I hypothesize that, in general, an increase in campaign ads will be a strong predictor for how well a candidate does in that area.

This question is important because politicians invest a lot of money into their campaign ads in order to increase their voter share. According to NPR, in 2016, presidential candidates were predicted to spend $4.4 billion (source: https://www.npr.org/sections/itsallpolitics/2015/08/19/432759311/2016-campaign-tv-ad-spending). However, do these campaign ads actually affect their election results? 

## Data Section
My sample is composed of two sources. 

The explanatory variable is proportion of Hillary Clinton campaign ads in a region. The outcome variable is the proportion of the vote that Hillary Clinton won. The explanatory variable was captured by the Political TV Ad Archive. It is comprised of campaign ads from the primary period through the 2016 election in regions across the United States. I sorted the data set to only include pro-presidential campaign ads. 

The dependent variable is the final election votes from the regions that the ad campaigns ran in. The final election votes were compiled from various Green Pages for the relevant districts. 

If I observe a strong correlation between campaign ads and election results, I will have support for my hypothesis. 

My study is cross-sectional—observing the effect of voting after exposing to TV ads. It is a census -- taking all the campaign ads from a district and all the votes from that district into consideration. 

```{r, echo=FALSE}
library(tidyverse)
library(dplyr)
```


```{r, echo=FALSE}
gov50ads <- read_csv("2016ads.csv", show_col_types = FALSE)
```

```{r, echo=FALSE}
gov50ads <- gov50ads |>
  filter(location != "Greenville-Spartanburg, SC/Asheville-Anderson, NC",
         location != "Washington, DC/Hagerstown, MD",
         location != "Boston, MA/Manchester, NH") |>
  mutate(state = case_when(
    location == "Las Vegas, NV" ~ "NV",
    location == "Tampa-St. Petersburg, FL" | location == "Orlando-Daytona Beach-Melbourne, FL" | location == "Miami-Fort Lauderdale, FL" ~ "FL",
    location == "Denver, CO" | location == "Colorado Springs-Pueblo, CO" ~ "CO",
    location == "Ceder Rapids-Waterloo-Iowa City-Dublin, Iowa" | location == "Sioux City, Iowa" | location == "Des Moines-Ames, Iowa" ~ "IA",
    location == "Charlotte, NC" | location == "Raleigh-Durham-Fayetteville,  NC" | location == "Norfolk-Portsmouth-Newport News, NC" ~ "NC",
    location == "Cincinnati, OH" | location == "Cleveland, Ohio" ~ "OH",
    location == "Columbia, SC" ~ "SC",
    location == "Milwaukee, WI" ~ "WI",
    location == "New York City, NY" ~ "NY",
    location == "Philadelphia, PA" ~ "PA",
    location == "Phoenix-Prescott, AZ" ~ "AZ",
    location == "Reno, NV" ~ "NV",
    location == "Roanoke-Lynchburg, VA" ~ "VA",
    location == "San Francisco-Oakland-San Jose, CA" ~ "CA"
      )) |>
  filter(race == "PRES",
         message == "pro",
         type == "campaign")
```

```{r, echo=FALSE}
gov50ads <- gov50ads |> 
  mutate(Hillad = if_else(str_detect(candidates, "Hillary Clinton"), 1, 0),
         Hillad = as.integer(Hillad))

```

### Data Wrangling

#### Political TV Ad Archive Data Set
The Political TV Ad Archive data set is comprised of 375,350 ads between October 2015 and December 2016. I filtered the ads to only include pro-presidential campaign ads.

The data set included ads from regions around the United States. Political TV Ad Archive stated that they selected these regions to create a historically, geographically, and politically diverse combination. For simplicity sake, I renamed the regions to their encompassing state (e.g., Tampa-St.Petersburg, Florida ads were renamed "FL"). I filtered out data that included multiple states (e.g., Washington, DC/Hagerstown, MD). 

After filtering, there were 57,800 ads in the data set.

#### Election Result Data Set
As mentioned earlier, the election results were comprised from the Green Pages. I only included election results from the districts mentioned in the Political TV Ad Archive Data Set (e.g., I did not take all the election results from South Carolina, only from the county that included Columbia). 

### Plots Summarizing Dependent Variable
```{r, echo=FALSE}
gov50ads |> 
  group_by(candidates) |>
  summarize(n = n()) |>
  mutate(prop = n / sum(n)) |>
  drop_na() |>
  slice_max(n, n = 6) |>
  ggplot(mapping = aes(x = candidates,
                       y = prop)) + 
  geom_col() + 
  labs(
    title = "Proportion of Pro-Campaign Ads Run from Oct 2015 - Dec 2016",
    y = "Proportion of Campaign Ads",
    x = "Candidate",
    caption = "Source = Political TV Ad Archive")
```

```{r, echo=FALSE}
gov50ads |>
  group_by(state) |>
  summarize(mean_Hillad = mean((Hillad), na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = mean_Hillad)) +
  geom_col() + 
  labs(title = "Proportion of Campaign Ads Belonging to Hillary Clinton in Different Districts",
       subtitle = "From October 2015 - December 2016",
       y = "Proportion of Ads Belonging to Clinton",
       x = "State",
       caption = "Source = Political TV Ad Archive")
```

## Results 

```{r, echo=FALSE}
presidential <- read_csv("gov50presidential.csv", show_col_types = FALSE)
presidential <- presidential |> 
  rename("state" = "...1") |>
  group_by(state) |>
  summarize(perc_Hillaryvote = mean(`Hillary Clinton`))
```

```{r, echo=FALSE}
join_ads <- gov50ads |> 
  group_by(state) |> 
  summarize(perc_Hillaryad = mean(Hillad)) |>
  left_join(presidential, by = "state") |> 
  mutate(`Hillary Clinton` = perc_Hillaryad * 100) 
```

```{r, echo=FALSE}
join_ads |> 
  ggplot(mapping = aes(x = `Hillary Clinton`,
                       y = perc_Hillaryvote)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm") + 
  labs(title = "Hillary Clinton 2016 Presidential Campaign Ads vs Election Results",
       y = "Clinton Percentage of 2016 Presidential Election",
       x = "Clinton Percentage of Campaign Ads", 
       caption = "Source: Political TV Ad Archive and Green Pages")
```

```{r, echo=FALSE}
join_lm <- join_ads |>
  lm(formula = `perc_Hillaryvote` ~ `Hillary Clinton`) 
```

```{r, echo=FALSE}
modelsummary::modelsummary(join_lm,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           title = "Linear Regression of Hillary Clinton Share of Ads and Votes")
```

### RESULT SUMMARY: 

In the scatter plot, we can see that as Hillary Clinton's proportion of campaign ads increased, her share of the vote actually decreased. According to the line of best fit model, the intercept is 79.88. That means that when Clinton ads make up 0% of the campaign ads, she has 77.88% of the vote for that district. The slope is -0.405. That is, for every increase in 1 percentage point of the campaign ads that Clinton's ads make up, her share of the vote decreases by 0.405 percentage points. 

Based on this study, we find that increasing candidate's ads make up of campaign ads are actually correlated with a decrease in share of election votes. 

The survey is not causal. It was an observational study, not a randomized experimental design. Additionally, share of campaign ads and election votes might be related. That is, a district with negative pre-existing notions of Hillary Clinton may be targeted with a greater percentage of ads in order to improve that perception. 

It is also not statistically significant. The p-value for the slope is 0.102. As this p-value is greater than the standard alpha value of 0.05, we fail to reject the null hypothesis that increasing percentage of campaign ads increase percentage of election votes. 

## Conclusion

My original hypothesis that increasing share of election ads would increase share of election votes was incorrect. I do not find support for my hypothesis. 

### LIMITATIONS / IMPROVEMENT:
One limitation of this study is that the 2016 presidential election was already extremely contentious. As such, it may not be the best study for the efficacy of campaign ads. A better study subject might be an election that is not as polarized/sensationalized and a lower-scale one (such as gubernatorial or congressional, although these seats can also be just as drama-ridden as presidential). 

Another limitation of this study was that we solely looked at the election results for the impact. A better study would have assessed continuously throughout the campaign how people's perceptions of the election may have changed with changes in share of campaign ads. 